\subsection{Introduction}

A PBX is only really useful if you can get calls into it.  Of course, you
can use Asterisk with VoIP calls (SIP, H.323, IAX, etc.), but you can also 
talk to the real PSTN through various cards.

Supported Hardware is divided into two general groups:  DAHDI devices and 
non-DAHDI devices.  The DAHDI compatible hardware supports pseudo-TDM 
conferencing and all call features through chan\_dahdi, whereas non-DAHDI 
compatible hardware may have different features.

\subsection{DAHDI compatible hardware}

\begin{itemize}
\item Digium, Inc. (Primary Developer of Asterisk)
      \url{http://www.digium.com}
  \begin{itemize}
     \item Analog Interfaces
     \begin{itemize}
       \item TDM400P - The TDM400P is a half-length PCI 2.2-compliant card that supports FXS and FXO station interfaces for connecting analog telephones and analog POTS lines through a PC.
       \item TDM800P - The TDM800P is a half-length PCI 2.2-compliant, 8 port card using Digium's VoiceBus technology that supports FXS and FXO station interfaces for connecting analog telephones and analog POTS lines through a PC.
       \item TDM2400P - The TDM2400P is a full-length PCI 2.2-compliant card for connecting analog telephones and analog POTS lines through a PC. It supports a combination of up to 6 FXS and/or FXO modules for a total of 24 lines.
     \end{itemize}
     \item Digital Interfaces
     \begin{itemize}
       \item TE412P - The TE412P offers an on-board DSP-based echo cancellation module. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE410P - The TE410P improves performance and scalability through bus mastering architecture. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE407P - The TE407P offers an on-board DSP-based echo cancellation module. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE405P - The TE405P improves performance and scalability through bus mastering architecture. It supports both E1, T1, J1 environments and is selectable on a per-card or per-port basis.
       \item TE212P - The TE212P offers an on-board DSP-based echo cancellation module. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE210P - The TE210P improves performance and scalability through bus mastering architecture. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE207P - The TE207P offers an on-board DSP-based echo cancellation module. It supports E1, T1, and J1 environments and is selectable on a per-card or per-port basis.
       \item TE205P - The TE205P improves performance and scalability through bus mastering architecture. It supports both E1 and T1/J1 environments and is selectable on a per-card or per-port basis.
       \item TE120P - The TE120P is a single span, selectable T1, E1, or J1 card and utilizes Digium's VoiceBus\texttrademark technology. It supports both voice and data modes.       
       \item TE110P - The TE110P brings a high-performance, cost-effective, and flexible single span togglable T1, E1, J1 interface to the Digium line-up of telephony interface devices.
     \end{itemize}
  \end{itemize}
\end{itemize}

\subsection{Non-DAHDI compatible hardware}

\begin{itemize}
  \item QuickNet, Inc. 
        \url{http://www.quicknet.net}
  \begin{itemize}
    \item Internet PhoneJack - Single FXS interface.  Supports Linux telephony
          interface.  DSP compression built-in.

    \item Internet LineJack - Single FXS or FXO interface.  Supports Linux 
          telephony interface.
  \end{itemize}
\end{itemize}

\subsection{mISDN compatible hardware}

mISDN homepage:  \url{http://www.misdn.org/}

Any adapter with an mISDN driver should be compatible with
chan\_misdn. See the mISDN section for more information.

\begin{itemize}
  \item Digium, Inc. (Primary Developer of Asterisk) 
  \url{http://www.digium.com}
  \begin{itemize}
    \item B410P - 4 Port BRI card (TE/NT)
   \end{itemize}
\end{itemize}

\begin{itemize}
  \item beroNet 
  \url{http://www.beronet.com}
  \begin{itemize}
    \item BN4S0 - 4 Port BRI card (TE/NT)

    \item BN8S0 - 8 Port BRI card (TE/NT)

    \item Billion Card - Single Port BRI card (TE (/NT with crossed cable))
   \end{itemize}
\end{itemize}

\subsection{Miscellaneous other interfaces}

\begin{itemize}
  \item Digium, Inc. (Primary Developer of Asterisk)
  \begin{itemize}
    \item TC400B - The TC400B is a half-length, low-profile PCI 2.2-compliant card for transforming complex VoIP codecs (G.729) into simple codecs.
  \end{itemize}

  \item ALSA
        \url{http://www.alsa-project.org}
  \begin{itemize}
    \item Any ALSA compatible full-duplex sound card
  \end{itemize}

  \item OSS
        \url{http://www.opensound.com}
  \begin{itemize}
    \item Any OSS compatible full-duplex sound card
  \end{itemize}
\end{itemize}
